---
title: "Modeling pragmatic inferences triggered by informational redundancy (additional models)"
output: 
  html_document:
    self_contained: false
    toc: true
    toc_float: true
    pandoc_args: --mathjax
link-citations: true
bibliography: bibliography.bib
csl: apa.csl
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rwebppl)
library(tidyverse)
library(fitdistrplus)
library(gridExtra)
library(scales)
library(grid)

data <- read_csv("data/before_after_ratings.csv") %>% mutate_at(vars(id,story:experiment,Qtype,condition:eligible), factor) %>% filter(eligible=="True")

prior_ratings <- data %>% filter(condition=="before" & Qtype %in% c("predictable","unpredictable") & world %in% c("typical","wonky"))

prior_typ_pred <- prior_ratings %>% filter(world=="typical" & Qtype=="predictable")
prior_wonk_pred <- prior_ratings %>% filter(world=="wonky" & Qtype=="predictable")
prior_typ_unpred <- prior_ratings %>% filter(world=="typical" & Qtype=="unpredictable")
prior_wonk_unpred <- prior_ratings %>% filter(world=="wonky" & Qtype=="unpredictable")

posterior_ratings <- data %>% filter(condition=="after" & as.character(Qtype) == as.character(activity) & world %in% c("typical","wonky"))

posterior_typ_pred <- posterior_ratings %>% filter(world=="typical" & activity=="predictable")
posterior_wonk_pred <- posterior_ratings %>% filter(world=="wonky" & activity=="predictable")
posterior_typ_unpred <- posterior_ratings %>% filter(world=="typical" & activity=="unpredictable")
posterior_wonk_unpred <- posterior_ratings %>% filter(world=="wonky" & activity=="unpredictable")

posterior_typ_pred_empty <- data %>% filter(condition=="after" & Qtype == "predictable" & world == "typical" & activity=="unpredictable")

posterior_wonk_pred_empty <- data %>% filter(condition=="after" & Qtype == "predictable" & world == "wonky" & activity=="unpredictable")

posterior_typ_unpred_empty <- data %>% filter(condition=="after" & Qtype == "unpredictable" & world == "typical" & activity=="predictable")

posterior_wonk_unpred_empty <- data %>% filter(condition=="after" & Qtype == "unpredictable" & world == "wonky" & activity=="predictable")

rsa_sourcecode <- paste(readLines("models/rsa.webppl"), collapse="\n")
hrsa_sourcecode <- paste(readLines("models/hrsa.webppl"), collapse="\n")
noisy_hrsa_sourcecode <- paste(readLines("models/noisy_hrsa.webppl"), collapse="\n")
nonhab_noisy_hrsa_sourcecode <- paste(readLines("models/nonhab_noisy_hrsa.webppl"), collapse="\n")

########

scale_ratings <- function(rating) {
  (rating/100 - min(rating/100) + 0.001) / (max(rating/100) - min(rating/100) + 0.002)
}

dist_one <- function(result,title) {
  ggplot(result, aes(x=habituality)) + geom_density(fill="#E69F00") + ylim(0,15) + ggtitle(title)
}

dist_both <- function(result,title,levels) {
  result$state <- factor(result$state, levels=levels)
  ggplot(result, aes(x=habituality, fill=state)) + geom_density(alpha=0.5) + scale_fill_manual(values=c("#E69F00", "#999999", "#56B4E9"), drop=FALSE) + ylim(0,32.5) + ggtitle(title)
}

results <- function(result, title, levels) {
  result <- rename(result, state=support)
  table <- tableGrob(result)
  result$state <- factor(result$state, levels=levels)
  plot <- ggplot(result, aes(x=state, y=prob)) +   geom_bar(stat="identity") + scale_y_continuous(labels=percent, name="percent", limits = c(0,1)) + scale_x_discrete(drop=FALSE)
  grid.arrange(top=textGrob(title, gp=gpar(fontsize=13.2)), table, plot, ncol=2)
}

prob_state <- function(result) {
  result %>% dplyr::select(support = state) %>% group_by(support) %>% tally() %>% mutate(prob = n / sum(n)) %>% dplyr::select(support,prob)
}

prob_habit <- function(result) {
  result %>% dplyr::select(habituality)
}

prob_both <- function(result) {
  result %>% spread(Parameter, value) %>% dplyr::select(-Iteration, -Chain) %>% mutate_at("habituality", as.numeric)
}
```

This page demonstrates the performance of the noisy channel hRSA with lower-*habituality* utterances.

## Empirical priors

### Wonky context - habitual activity

**Context: **"*John is typically broke, and doesn't usually pay when he goes to the grocery store.*"  
**Question: **"*How often do you think John usually pays the cashier, when grocery shopping?*"

In this case, the distribution of responses suggests that participants differ in whether they adjust their estimates for *activity habituality* when a context is 'wonky' (and either clearly states, or suggests, that the activity is unpredictable), although a large number shift their estimates considerably.  This likely indicates a relative insensitivity to the context, particularly where it does not state quite as bluntly that the activity usually doesn't occur.  Overall, however, comprehenders who see the informationally redundant utterance are relatively likely to conclude that *John* is *not* a habitual cashier-payer.

<!-- **CONSIDER INCLUDING NORMING RATINGS FROM SIMPLER CONTEXT SENTENCES, BUT MAKES MORE DIFFICULT TO COMPARE EMPIRICAL RESULTS** -->

To note, for this set of ratings, and those following, there is a clear 'bump' mid-distribution, suggesting that participants have a tendency to click right in the middle of the slider when the rating roughly approximates 50%.  This does harm distribution fit somewhat.

```{r wonk pred plot, echo=FALSE, fig.height=4}
# Histogram with density plot
ggplot(prior_wonk_pred, aes(x=rating)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white", bins=50) +
  geom_density(alpha=.2, fill="#FF6666") +
  geom_vline(aes(xintercept=mean(rating)), color="black", linetype="dashed", size=1)
```

```{r wonk pred fit, cache=TRUE, echo=FALSE}
# scale to remove 0 and 1 values (add/subtract 0.001 from edges)
prior_wonk_pred_scaled <- scale_ratings(prior_wonk_pred$rating)

# fit beta distribution by maximum likelihood estimation
fit.prior_wonk_pred <- fitdist(prior_wonk_pred_scaled, "beta", method="mle")
summary(fit.prior_wonk_pred)
plot(fit.prior_wonk_pred)

js.prior_pred <- paste0("var beta_high_a = 2.19602561493963","\n","var beta_high_b = 0.410857186535822","\n","var beta_low_a = ",fit.prior_wonk_pred$estimate[1],"\n","var beta_low_b = ",fit.prior_wonk_pred$estimate[2],"\n")
```

### Typical context - non-habitual activity

**Context: **"*John often goes to the grocery store around the corner from his apartment.*"  
**Question: **"*How often do you think John usually gets apples, when grocery shopping?*"

For this set of ratings and the following, it is clear that participants take apple-bying to be a moderately habitual to relatively non-habitual activity.

```{r typ unpred plot, echo=FALSE, fig.height=4}
# Histogram with density plot
ggplot(prior_typ_unpred, aes(x=rating)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white", bins=50) +
  geom_density(alpha=.2, fill="#FF6666") +
  geom_vline(aes(xintercept=mean(rating)), color="black", linetype="dashed", size=1)
```

```{r typ unpred fit, cache=TRUE, echo=FALSE}
# scale to remove 0 and 1 values (add/subtract 0.001 from edges)
prior_typ_unpred_scaled <- scale_ratings(prior_typ_unpred$rating)

# fit beta distribution by maximum likelihood estimation
fit.prior_typ_unpred <- fitdist(prior_typ_unpred_scaled, "beta", method="mle")
summary(fit.prior_typ_unpred)
plot(fit.prior_typ_unpred)
```

### Wonky context - non-habitual activity

**Context: **"*John is typically broke, and doesn't usually pay when he goes to the grocery store.*"  
**Question: **"*How often do you think John usually gets apples, when grocery shopping?*"

Here, it appears that comprehenders are only slightly affected by the *wonky* context - *activity habituality* estimates do decrease slightly, but likely as a result of a general suspicion that *John* does not engage in very typical grocery shopping.

```{r wonk unpred plot, echo=FALSE, fig.height=4}
# Histogram with density plot
ggplot(prior_wonk_unpred, aes(x=rating)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white", bins=50) +
  geom_density(alpha=.2, fill="#FF6666") +
  geom_vline(aes(xintercept=mean(rating)), color="black", linetype="dashed", size=1)
```

```{r wonk unpred fit, cache=TRUE, echo=FALSE}
# scale to remove 0 and 1 values (add/subtract 0.001 from edges)
prior_wonk_unpred_scaled <- scale_ratings(prior_wonk_unpred$rating)

# fit beta distribution by maximum likelihood estimation
fit.prior_wonk_unpred <- fitdist(prior_wonk_unpred_scaled, "beta", method="mle")
summary(fit.prior_wonk_unpred)
plot(fit.prior_wonk_unpred)


js.prior_pred_nonhab <- paste0("var beta_high_a = ",fit.prior_typ_unpred$estimate[1],"\n","var beta_high_b = ",fit.prior_typ_unpred$estimate[1],"\n","var beta_low_a = ",fit.prior_wonk_unpred$estimate[1],"\n","var beta_low_b = ",fit.prior_wonk_unpred$estimate[2],"\n")
```

## Wonky habitual (non-payer): "He paid the cashier!"

```{r wonky noisy hRSA model, echo=FALSE, cache=TRUE}
base_model <- paste0(js.prior_pred, "\nvar world_type = \"wonky\"\n", noisy_hrsa_sourcecode)

l0_95 <- webppl(paste(base_model, "literalListener(\"(...)\",0.95)", sep="\n"))
l0_50 <- webppl(paste(base_model, "literalListener(\"(...)\",0.5)", sep="\n"))
l0_5 <- webppl(paste(base_model, "literalListener(\"(...)\",0.05)", sep="\n"))

s_95 <- webppl(paste(base_model, "speaker(\"happened\",0.95)", sep="\n"))
s_50 <- webppl(paste(base_model, "speaker(\"happened\",0.5)", sep="\n"))
s_5 <- webppl(paste(base_model, "speaker(\"happened\",0.05)", sep="\n"))

l1_empty <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"(...)\",\"both\")", sep="\n"))))
l1_plain <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"plain\",\"both\")", sep="\n"))))
l1_exclamation <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"exclamation\",\"both\")", sep="\n"))))
l1_ohyeah <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"oh yeah\",\"both\")", sep="\n"))))

l1_empty_habit <- prob_habit(l1_empty)
l1_plain_habit <- prob_habit(l1_plain)
l1_exclamation_habit <- prob_habit(l1_exclamation)
l1_ohyeah_habit <- prob_habit(l1_ohyeah)

l1_empty_state <- prob_state(l1_empty)
l1_plain_state <- prob_state(l1_plain)
l1_exclamation_state <- prob_state(l1_exclamation)
l1_ohyeah_state <- prob_state(l1_ohyeah)
```

```{r wonky noisy hRSA plots 3, fig.height = 2, echo=FALSE}
utterances <- c("(...)","plain","exclamation","oh yeah")
states <- c("happened","didn't happen")

dist_one(l1_empty_habit,"Pragmatic Listener: \"(...)\" (`habituality` only)")
dist_one(l1_plain_habit,"Pragmatic Listener: \"John paid the cashier.\" (`habituality` only)")
dist_one(l1_exclamation_habit,"Pragmatic Listener: \"John paid the cashier!\" (`habituality` only)")
dist_one(l1_ohyeah_habit,"Pragmatic Listener: \"Oh yeah, and John paid the cashier.\" (`habituality` only)")

results(l1_empty_state,"Pragmatic Listener: \"(...)\" (`state` only)", states)
results(l1_plain_state,"Pragmatic Listener: \"John paid the cashier.\" (`state` only)", states)
results(l1_exclamation_state,"Pragmatic Listener: \"John paid the cashier!\" (`state` only)", states)
results(l1_ohyeah_state,"Pragmatic Listener: \"Oh yeah, and John paid the cashier.\" (`state` only)", states)

dist_both(l1_empty,"Pragmatic Listener: \"(...)\"", states)
dist_both(l1_plain,"Pragmatic Listener: \"John paid the cashier.\"", states)
dist_both(l1_exclamation,"Pragmatic Listener: \"John paid the cashier!\"", states)
dist_both(l1_ohyeah,"Pragmatic Listener: \"Oh yeah, and John paid the cashier.\"", states)
```

### Comparison to empirical results

```{r wonky empirical results 1, fig.height = 6, echo=FALSE}
data_empty <- tibble(rating = posterior_wonk_pred_empty$rating/100)
data_plain <- (filter(posterior_wonk_pred, experiment=="period") %>% dplyr::select(rating))/100
data_exclamation <- (filter(posterior_wonk_pred, experiment=="exclamation") %>% dplyr::select(rating))/100
data_ohyeah <- (filter(posterior_wonk_pred, experiment=="ohyeah") %>% dplyr::select(rating))/100

# l1_empty_habit
# l1_plain_habit
# l1_exclamation_habit
# l1_ohyeah_habit

empirical <- nrow(data_empty) + nrow(posterior_wonk_pred)
predicted <- nrow(l1_empty_habit)*4

type <- c(rep("Empirical",nrow(data_empty)), rep("Predicted",nrow(l1_empty_habit)), rep("Empirical",nrow(data_plain)), rep("Predicted",nrow(l1_plain_habit)), rep("Empirical",nrow(data_exclamation)), rep("Predicted",nrow(l1_exclamation_habit)), rep("Empirical",nrow(data_ohyeah)), rep("Predicted",nrow(l1_ohyeah_habit)))

experiment <- c(rep("\"(...)\"",nrow(data_empty)+nrow(l1_empty_habit)), rep("\"John paid the\ncashier.\"",nrow(data_plain)+nrow(l1_plain_habit)), rep("\"John paid the\ncashier!\"",nrow(data_exclamation)+nrow(l1_exclamation_habit)), rep("\"Oh yeah, and John\npaid the cashier.\"",nrow(data_ohyeah)+nrow(l1_ohyeah_habit)))

comparison <- tibble(type = type, experiment = experiment, habituality = c(data_empty$rating, l1_empty_habit$habituality,data_plain$rating,l1_plain_habit$habituality,data_exclamation$rating,l1_exclamation_habit$habituality,data_ohyeah$rating,l1_ohyeah_habit$habituality)) %>% mutate_if(is.character, as.factor)

comparison$experiment <- factor(comparison$experiment, levels=c("\"(...)\"","\"John paid the\ncashier.\"","\"John paid the\ncashier!\"","\"Oh yeah, and John\npaid the cashier.\""))

ggplot(comparison, aes(x=habituality)) + geom_density(fill="#E69F00") + facet_grid(experiment ~ type) + ggtitle("Empirical vs. Predicted Probability Densities")
```

```{r wonky empirical results 2, fig.height = 4, echo=FALSE}
comparison %>% group_by(type,experiment) %>% summarize(habituality = mean(habituality)) %>% ggplot(aes(x=type, y=habituality, fill=type)) + geom_bar(stat="identity") + facet_grid(. ~ experiment) + scale_y_continuous(labels = percent_format(), limits = c(0,1)) + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) + ggtitle("Empirical vs. Predicted Habituality Means")
```

## Ordinary non-habitual (presumed payer): "He got some apples!"

```{r ordinary nonhab noisy hRSA model, echo=FALSE, cache=TRUE}
base_model <- paste0(js.prior_pred_nonhab, "\nvar world_type = \"ordinary\"\n", nonhab_noisy_hrsa_sourcecode)

l0_95 <- webppl(paste(base_model, "literalListener(\"(...)\",0.95)", sep="\n"))
l0_50 <- webppl(paste(base_model, "literalListener(\"(...)\",0.5)", sep="\n"))
l0_5 <- webppl(paste(base_model, "literalListener(\"(...)\",0.05)", sep="\n"))

s_95 <- webppl(paste(base_model, "speaker(\"happened\",0.95)", sep="\n"))
s_50 <- webppl(paste(base_model, "speaker(\"happened\",0.5)", sep="\n"))
s_5 <- webppl(paste(base_model, "speaker(\"happened\",0.05)", sep="\n"))

l1_empty <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"(...)\",\"both\")", sep="\n"))))
l1_plain <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"plain\",\"both\")", sep="\n"))))
l1_exclamation <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"exclamation\",\"both\")", sep="\n"))))
l1_ohyeah <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"oh yeah\",\"both\")", sep="\n"))))

l1_empty_habit <- prob_habit(l1_empty)
l1_plain_habit <- prob_habit(l1_plain)
l1_exclamation_habit <- prob_habit(l1_exclamation)
l1_ohyeah_habit <- prob_habit(l1_ohyeah)

l1_empty_state <- prob_state(l1_empty)
l1_plain_state <- prob_state(l1_plain)
l1_exclamation_state <- prob_state(l1_exclamation)
l1_ohyeah_state <- prob_state(l1_ohyeah)
```


```{r ordinary nonhab noisy hRSA plots 3, fig.height = 2, echo=FALSE}
dist_one(l1_empty_habit,"Pragmatic Listener: \"(...)\" (`habituality` only)")
dist_one(l1_plain_habit,"Pragmatic Listener: \"John got some apples.\" (`habituality` only)")
dist_one(l1_exclamation_habit,"Pragmatic Listener: \"John got some apples!\" (`habituality` only)")
dist_one(l1_ohyeah_habit,"Pragmatic Listener: \"Oh yeah, and John got some apples.\" (`habituality` only)")

results(l1_empty_state,"Pragmatic Listener: \"(...)\" (`state` only)", states)
results(l1_plain_state,"Pragmatic Listener: \"John got some apples.\" (`state` only)", states)
results(l1_exclamation_state,"Pragmatic Listener: \"John got some apples!\" (`state` only)", states)
results(l1_ohyeah_state,"Pragmatic Listener: \"Oh yeah, and John got some apples.\" (`state` only)", states)

dist_both(l1_empty,"Pragmatic Listener: \"(...)\"", states)
dist_both(l1_plain,"Pragmatic Listener: \"John got some apples.\"", states)
dist_both(l1_exclamation,"Pragmatic Listener: \"John got some apples!\"", states)
dist_both(l1_ohyeah,"Pragmatic Listener: \"Oh yeah, and John got some apples.\"", states)
```

### Comparison to empirical results

```{r ordinary nonhab empirical results 1, fig.height = 6, echo=FALSE}
data_empty <- tibble(rating = posterior_typ_unpred_empty$rating/100)
data_plain <- (filter(posterior_typ_unpred, experiment=="period") %>% dplyr::select(rating))/100
data_exclamation <- (filter(posterior_typ_unpred, experiment=="exclamation") %>% dplyr::select(rating))/100
data_ohyeah <- (filter(posterior_typ_unpred, experiment=="ohyeah") %>% dplyr::select(rating))/100

# l1_empty_habit
# l1_plain_habit
# l1_exclamation_habit
# l1_ohyeah_habit

empirical <- nrow(data_empty) + nrow(posterior_typ_unpred)
predicted <- nrow(l1_empty_habit)*4

type <- c(rep("Empirical",nrow(data_empty)), rep("Predicted",nrow(l1_empty_habit)), rep("Empirical",nrow(data_plain)), rep("Predicted",nrow(l1_plain_habit)), rep("Empirical",nrow(data_exclamation)), rep("Predicted",nrow(l1_exclamation_habit)), rep("Empirical",nrow(data_ohyeah)), rep("Predicted",nrow(l1_ohyeah_habit)))

experiment <- c(rep("\"(...)\"",nrow(data_empty)+nrow(l1_empty_habit)), rep("\"John paid the\ncashier.\"",nrow(data_plain)+nrow(l1_plain_habit)), rep("\"John paid the\ncashier!\"",nrow(data_exclamation)+nrow(l1_exclamation_habit)), rep("\"Oh yeah, and John\ngot some apples.\"",nrow(data_ohyeah)+nrow(l1_ohyeah_habit)))

comparison <- tibble(type = type, experiment = experiment, habituality = c(data_empty$rating, l1_empty_habit$habituality,data_plain$rating,l1_plain_habit$habituality,data_exclamation$rating,l1_exclamation_habit$habituality,data_ohyeah$rating,l1_ohyeah_habit$habituality)) %>% mutate_if(is.character, as.factor)

comparison$experiment <- factor(comparison$experiment, levels=c("\"(...)\"","\"John paid the\ncashier.\"","\"John paid the\ncashier!\"","\"Oh yeah, and John\ngot some apples.\""))

ggplot(comparison, aes(x=habituality)) + geom_density(fill="#E69F00") + facet_grid(experiment ~ type) + ggtitle("Empirical vs. Predicted Probability Densities")
```

```{r ordinary nonhab empirical results 2, fig.height = 4, echo=FALSE}
comparison %>% group_by(type,experiment) %>% summarize(habituality = mean(habituality)) %>% ggplot(aes(x=type, y=habituality, fill=type)) + geom_bar(stat="identity") + facet_grid(. ~ experiment) + scale_y_continuous(labels = percent_format(), limits = c(0,1)) + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) + ggtitle("Empirical vs. Predicted Habituality Means")
```


## Wonky non-habitual (non-payer): "He got some apples!"

```{r wonky nonhab noisy hRSA model, echo=FALSE, cache=TRUE}
base_model <- paste0(js.prior_pred_nonhab, "\nvar world_type = \"wonky\"\n", nonhab_noisy_hrsa_sourcecode)

l0_95 <- webppl(paste(base_model, "literalListener(\"(...)\",0.95)", sep="\n"))
l0_50 <- webppl(paste(base_model, "literalListener(\"(...)\",0.5)", sep="\n"))
l0_5 <- webppl(paste(base_model, "literalListener(\"(...)\",0.05)", sep="\n"))

s_95 <- webppl(paste(base_model, "speaker(\"happened\",0.95)", sep="\n"))
s_50 <- webppl(paste(base_model, "speaker(\"happened\",0.5)", sep="\n"))
s_5 <- webppl(paste(base_model, "speaker(\"happened\",0.05)", sep="\n"))

l1_empty <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"(...)\",\"both\")", sep="\n"))))
l1_plain <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"plain\",\"both\")", sep="\n"))))
l1_exclamation <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"exclamation\",\"both\")", sep="\n"))))
l1_ohyeah <- prob_both(as_tibble(webppl(paste(base_model, "pragmaticListener(\"oh yeah\",\"both\")", sep="\n"))))

l1_empty_habit <- prob_habit(l1_empty)
l1_plain_habit <- prob_habit(l1_plain)
l1_exclamation_habit <- prob_habit(l1_exclamation)
l1_ohyeah_habit <- prob_habit(l1_ohyeah)

l1_empty_state <- prob_state(l1_empty)
l1_plain_state <- prob_state(l1_plain)
l1_exclamation_state <- prob_state(l1_exclamation)
l1_ohyeah_state <- prob_state(l1_ohyeah)
```

```{r wonky nonhab noisy hRSA plots 3, fig.height = 2, echo=FALSE}
dist_one(l1_empty_habit,"Pragmatic Listener: \"(...)\" (`habituality` only)")
dist_one(l1_plain_habit,"Pragmatic Listener: \"John got some apples.\" (`habituality` only)")
dist_one(l1_exclamation_habit,"Pragmatic Listener: \"John got some apples!\" (`habituality` only)")
dist_one(l1_ohyeah_habit,"Pragmatic Listener: \"Oh yeah, and John got some apples.\" (`habituality` only)")

results(l1_empty_state,"Pragmatic Listener: \"(...)\" (`state` only)", states)
results(l1_plain_state,"Pragmatic Listener: \"John got some apples.\" (`state` only)", states)
results(l1_exclamation_state,"Pragmatic Listener: \"John got some apples!\" (`state` only)", states)
results(l1_ohyeah_state,"Pragmatic Listener: \"Oh yeah, and John got some apples.\" (`state` only)", states)

dist_both(l1_empty,"Pragmatic Listener: \"(...)\"", states)
dist_both(l1_plain,"Pragmatic Listener: \"John got some apples.\"", states)
dist_both(l1_exclamation,"Pragmatic Listener: \"John got some apples!\"", states)
dist_both(l1_ohyeah,"Pragmatic Listener: \"Oh yeah, and John got some apples.\"", states)
```

### Comparison to empirical results

```{r wonky nonhab empirical results 1, fig.height = 6, echo=FALSE}
data_empty <- tibble(rating = posterior_wonk_unpred_empty$rating/100)
data_plain <- (filter(posterior_wonk_unpred, experiment=="period") %>% dplyr::select(rating))/100
data_exclamation <- (filter(posterior_wonk_unpred, experiment=="exclamation") %>% dplyr::select(rating))/100
data_ohyeah <- (filter(posterior_wonk_unpred, experiment=="ohyeah") %>% dplyr::select(rating))/100

# l1_empty_habit
# l1_plain_habit
# l1_exclamation_habit
# l1_ohyeah_habit

empirical <- nrow(data_empty) + nrow(posterior_wonk_unpred)
predicted <- nrow(l1_empty_habit)*4

type <- c(rep("Empirical",nrow(data_empty)), rep("Predicted",nrow(l1_empty_habit)), rep("Empirical",nrow(data_plain)), rep("Predicted",nrow(l1_plain_habit)), rep("Empirical",nrow(data_exclamation)), rep("Predicted",nrow(l1_exclamation_habit)), rep("Empirical",nrow(data_ohyeah)), rep("Predicted",nrow(l1_ohyeah_habit)))

experiment <- c(rep("\"(...)\"",nrow(data_empty)+nrow(l1_empty_habit)), rep("\"John paid the\ncashier.\"",nrow(data_plain)+nrow(l1_plain_habit)), rep("\"John paid the\ncashier!\"",nrow(data_exclamation)+nrow(l1_exclamation_habit)), rep("\"Oh yeah, and John\ngot some apples.\"",nrow(data_ohyeah)+nrow(l1_ohyeah_habit)))

comparison <- tibble(type = type, experiment = experiment, habituality = c(data_empty$rating, l1_empty_habit$habituality,data_plain$rating,l1_plain_habit$habituality,data_exclamation$rating,l1_exclamation_habit$habituality,data_ohyeah$rating,l1_ohyeah_habit$habituality)) %>% mutate_if(is.character, as.factor)

comparison$experiment <- factor(comparison$experiment, levels=c("\"(...)\"","\"John paid the\ncashier.\"","\"John paid the\ncashier!\"","\"Oh yeah, and John\ngot some apples.\""))

ggplot(comparison, aes(x=habituality)) + geom_density(fill="#E69F00") + facet_grid(experiment ~ type) + ggtitle("Empirical vs. Predicted Probability Densities")
```

```{r wonky nonhab empirical results 2, fig.height = 4, echo=FALSE}
comparison %>% group_by(type,experiment) %>% summarize(habituality = mean(habituality)) %>% ggplot(aes(x=type, y=habituality, fill=type)) + geom_bar(stat="identity") + facet_grid(. ~ experiment) + scale_y_continuous(labels = percent_format(), limits = c(0,1)) + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) + ggtitle("Empirical vs. Predicted Habituality Means")
```
